---
title: "Sedproxy: Simulation of Sediment Archived Climate Proxy Records."
output: 
  html_document: 
    keep_md: yes
---

------------------------------

## Introduction

`sedproxy` provides a forward model for sediment archived climate proxies. It is based on work described in Laepple and Huybers (2013). A manuscript is in preparation, Dolman and Laepple (in prep.), which will more fully describe the forward model and its applications. Please contact Dr Andrew Dolman <<andrew.dolman@awi.de>>, or Dr Thomas Laepple <<tlaepple@awi.de>>, at the Alfred-Wegener-Institute, Helmholtz Centre for Polar and Marine Research, Germany, for more information.

 
## Installation

**sedproxy** can be installed directly from bitbucket

```{r, eval=FALSE}
if (!require("devtools")) {
  install.packages("devtools")
}

devtools::install_bitbucket("ecus/sedproxy")
```

## Shiny app

A Shiny app provides a graphical user interface to the main `sedproxy` function `ClimToProxyClim`. Currently this app forward models a proxy from a randomly generated input climate and serves only to illustrate the functionality of `sedproxy`. The user can alter the parameters of the forward model and observe the results graphically. In future versions it may be possible to load input climates and other input parameters from a file and to save the output.

To run the Shiny app enter the following in the R terminal:

```{r run_shiny, eval=FALSE}
library(sedproxy)
ShinySedproxy()
```

Below are some brief examples of how to run `sedproxy` from a script or the command line.

## Example data

`sedproxy` includes example data for a single sediment core and location: core number 41 in the Shakun et al. (2012) compilation (`r sedproxy::N41.proxy.details$Core`, Rosenthal et al. 2003). The climate signal is taken from the [TraCE-21ka](http://www.cgd.ucar.edu/ccr/TraCE/) Simulation of Transient Climate Evolution over the last 21,000 years, using the grid cell closest to core `r sedproxy::N41.proxy.details$Core`. Seasonality of *G.ruber*, the Foraminifera for which test Mg/Ca ratios were measured, is taken from the model of Fraile et al (2008). Sediment accumulation rates were estimated from the depth and age data associated with core `r sedproxy::N41.proxy.details$Core`, with a minimum rate of 0.2 * the mean rate.


**The `r sedproxy::N41.proxy.details$Core` core**

```{r message = FALSE, warning = FALSE}
library(tidyverse)
library(knitr)
library(sedproxy)
```


```{r}
N41.proxy.details %>% 
  gather() %>% 
  kable(., format = "markdown")
```


**Input climate signal**

The first 5 rows:

```{r}
(N41.t21k.climate[1:5,]-273.15) %>% 
  kable(., format = "markdown", digits = 2)
```

**Actual proxy record**

Core `r sedproxy::N41.proxy.details$Core` (Rosenthal et al. 2003)

```{r}
kable(head(N41.proxy), format = "markdown")
```

*******

## Function `ClimToProxyClim`

`ClimToProxyClim` is the main function in package `sedproxy`. It simulates a sediment archived proxy from an assumed true climate signal, the sediment accumulation rate, seasonality of the encoding organism/process, and the number of samples per timepoint.


```{r , fig.show='hold', warning=FALSE}
set.seed(26052017)
clim.in <- N41.t21k.climate[nrow(N41.t21k.climate):1,] - 273.15

PFM <- ClimToProxyClim(clim.signal = clim.in,
                       timepoints = round(N41.proxy$Published.age),
                       proxy.calibration.type = "identity",
                       seas.prod = N41.G.ruber.seasonality,
                       sed.acc.rate = N41.proxy$Sed.acc.rate.m.yr,
                       meas.noise = 0.46, n.samples = 30,
                       n.replicates = 10)
```


```{r}
PFM$everything
```

**Simple plotting**

```{r default_plot, fig.width=6, fig.height=5.5}
PFM$everything %>% 
  PlotPFMs(max.replicates = 1)
```


**Plot 5 replicates of the final simulated proxy**

```{r plot_reps, fig.width=6, fig.height=5.5}
PFM$everything %>% 
  filter(stage == "simulated.proxy") %>% 
  PlotPFMs(., max.replicates = 5)
```


```{r MgCa, fig.show='hold', warning=FALSE, echo=FALSE, eval=FALSE}
# ### Proxy types
# 
# The initial input climate signal can be converted into "proxy units" if a `proxy.calibration.type` is specified. This simulates the Environment -> Sensor stage of the proxy system.

set.seed(26052017)
PFM_2 <- ClimToProxyClim(clim.signal = N41.t21k.climate[nrow(N41.t21k.climate):1,] - 273.15, 
                         timepoints = round(N41.proxy$Published.age),
                         proxy.calibration.type = "MgCa",
                         seas.prod = N41.G.ruber.seasonality,
                         sed.acc.rate = N41.proxy$Sed.acc.rate.m.yr,
                         meas.noise = 0.46, n.samples = 30,
                         n.replicates = 1)
```


```{r MgCa_plot, fig.width=6, fig.height=5.5, echo=FALSE, eval=FALSE}
# `PlotPFM` returns a ggplot2 object so it is easy to customize figures with new axes labels an such

PFM_2$everything %>% 
  PlotPFMs(.) +
  labs(y = "Mg/Ca")
```

```{r, echo=FALSE}
# If the simulated MgCa values are back-transformed into °C, they are not identical to the values obtained without initial transformation to MgCa ratios. This is because the relationship between MgCa and temperature is non-linear, so that the back-transformed mean of multiple MgCa ratios is not the same as the mean of the original temperatures that were encoded as MgCa. 


# a <- PFM$simulated.proxy$clim.signal.ann
# b <- PFM_2$simulated.proxy$clim.signal.ann
# d <- ProxyConversion(proxy.value = b, proxy.calibration.type = "MgCa")
# df <- (a-d)
# 
# wts <- N41.G.ruber.seasonality / sum(N41.G.ruber.seasonality)
# tmps <- (N41.t21k.climate[nrow(N41.t21k.climate):1,] - 273.15)[PFM$simulated.proxy$timepoints,]
# 
# var1 <- (apply(tmps, 1, function(x) var(x)))
# var2 <- (apply(tmps, 1, function(x) sum(((x - mean(x))^2 * wts))))
# 
# plot(var1, df, xlim = c(0, 5), ylim = c(0, -0.2))
# x <- 0:4
# 
# ## Approximation of the difference between the mean of temperature and the back-transformed mean of MgCa
# ## var * 1/2 * A * log(B)
# ## where MgCa = B * exp(A * T)
# 
# lines(x, x * 0.5 * 0.09 * log(0.38))
# abline(h = 0)
```







## Literature cited

Fraile, I., Schulz, M., Mulitza, S., & Kucera, M. (2008): Predicting the global distribution of planktonic foraminifera using a dynamic ecosystem model. Biogeosciences, 5: 891–911.

Laepple, T., & Huybers, P. (2013): Reconciling discrepancies between Uk37 and Mg/Ca reconstructions of Holocene marine temperature variability. Earth and Planetary Science Letters, 375: 418–429.

Rosenthal, Y., Oppo, D. W., & Linsley, B. K. (2003): The amplitude and phasing of climate change during the last deglaciation in the Sulu Sea, western equatorial Pacific. Geophys. Res. Lett., 30: 1428.

Shakun, J. D., Clark, P. U., He, F., Marcott, S. A., Mix, A. C., Liu, Z., Otto-Bliesner, B., Schmittner, A., & Bard, E. (2012): Global warming preceded by increasing carbon dioxide concentrations during the last deglaciation. Nature, 484: 49–54.


