---
title: "Using the Sedproxy Package"
author: "Vignette Author"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Using the Sedproxy Package}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

## function `BioturbateTimeseries`

An event occurs at timestep 100, its signal is "mixed in" to older sediment up to 1 * the bioturbation depth, and into future sediments.

### With 30 individuals sampled

```{r , fig.show='hold'}
library(sedproxy)
tss <- ts(c(rep(0, 100), 1, rep(0, 99)), 0:199)
timepoints <- c(time(tss))
bio_depth <- 0.10
acc_rate <- 0.02
NForams <- 30
set.seed(1)
result <- BioturbateTimeseries(tss, timepoints, bio_depth = bio_depth, acc_rate = acc_rate, n_samples = NForams)
plot(tss)
abline(v = 100, col = "Red")
tss2 <- ts(result$data, result$time)
plot(tss2)
abline(v = 100, col = "Red")
abline(v = 100 + bio_depth / acc_rate, col = "Blue")
```


### With infinite samples.

```{r , fig.show='hold'}
set.seed(1)
result <- BioturbateTimeseries(tss, timepoints, bio_depth = bio_depth, acc_rate = acc_rate, n_samples = Inf)
plot(tss)
abline(v = 100, col = "Red")
tss2 <- ts(result$data, result$time)
plot(tss2)
abline(v = 100, col = "Red")
abline(v = 100 + bio_depth / acc_rate, col = "Blue")
```


A sin wave represents the true process, bioturbation reduces the amplitude and shifts the signal. Sampling of 30 forams adds noise.

 
```{r}
tss<-ts(sin(seq(-pi, 9*pi, length.out = 1000)), start = 90, frequency = 1/100)
timepoints<-c(time(tss)[1:1000])

result <- BioturbateTimeseries(tss,timepoints, bio_depth = 0.1, acc_rate = 0.1, n_samples = 30)

plot(tss)
lines(data~time, data = result, col = "Red")
```

Sampling 300 forams.
 
```{r}
tss<-ts(sin(seq(-pi, 9*pi, length.out = 1000)), start = 0, frequency = 1/100)
timepoints<-c(time(tss)[1:1000])
result<-BioturbateTimeseries(tss,timepoints, bio_depth = 0.1, acc_rate = 0.1, n_samples = 300)
plot(tss)
lines(data~time, data = result, col = "Red")
```

Infinite sample.
 
```{r}
tss<-ts(sin(seq(-pi, 9*pi, length.out = 1000)), start = 0, frequency = 1/100)
timepoints<-c(time(tss)[1:1000])
result<-BioturbateTimeseries(tss,timepoints, bio_depth = 0.1, acc_rate = 0.1, n_samples = Inf)
plot(tss)
lines(data~time, data = result, col = "Red")
```

## function `SeasonalityBias`

```{r}
set.seed(1)
timepoints <- c(time(tss)[runif(10, 1, length(tss))])
SeasonalityBias(sst.clim = 1:12 * 0.1,
                                timepoints = timepoints,
                                proxy.clim = 12:1,
                                n_samples = Inf
                                )
```

```{r}
SeasonalityBias(sst.clim = 1:12 * 0.1,
                                timepoints = timepoints,
                                proxy.clim = 12:1,
                                n_samples = 30
                                )
```


## Combine bioturbation with seasonality bias

```{r}
tss<-ts(sin(seq(-pi, 9*pi, length.out = 1000)), start = 90, frequency = 1/100)
timepoints<-c(time(tss)[1:1000])
biowidth_timesteps=50
result <- BioturbateTimeseries(tss,timepoints, bio_depth = 0.1, acc_rate = 0.1, n_samples = 30)
plot(tss)
lines(data~time, data = result, col = "Red")

seas <- SeasonalityBias(sst.clim = 1:12 * 0.1,
                                timepoints = timepoints,
                                proxy.clim = 12:1,
                                n_samples = 30
                                )

result2 <- result
result2$data <- result2$data + seas$data
lines(data~time, data = result2, col = "Blue")
```


### Sample 100 time points at uniform random intervals

```{r}
tss<-ts(sin(seq(-pi, 9*pi, length.out = 1000)), start = 90, frequency = 1/100)
timepoints <- c(time(tss)[sort(runif(100, 1, length(tss)))])
result <- BioturbateTimeseries(tss, timepoints, bio_depth = 0.1, acc_rate = 0.1, n_samples = 30)
plot(tss)
lines(data~time, data = result, col = "Red")

seas <- SeasonalityBias(sst.clim = 1:12 * 0.1,
                                timepoints = timepoints,
                                proxy.clim = 12:1,
                                n_samples = 30
                                )

result2 <- result
result2$data <- result2$data + seas$data
lines(data~time, data = result2, col = "Blue")
```



Vignettes are long form documentation commonly included in packages. Because they are part of the distribution of the package, they need to be as compact as possible. The `html_vignette` output type provides a custom style sheet (and tweaks some options) to ensure that the resulting html is as small as possible. The `html_vignette` format:

- Never uses retina figures
- Has a smaller default figure size
- Uses a custom CSS stylesheet instead of the default Twitter Bootstrap style

## Vignette Info

Note the various macros within the `vignette` section of the metadata block above. These are required in order to instruct R how to build the vignette. Note that you should change the `title` field and the `\VignetteIndexEntry` to match the title of your vignette.

## Styles

The `html_vignette` template includes a basic CSS theme. To override this theme you can specify your own CSS in the document metadata as follows:

    output: 
      rmarkdown::html_vignette:
        css: mystyles.css

## Figures

The figure sizes have been customised so that you can easily put two images side-by-side. 

```{r, fig.show='hold'}
plot(1:10)
plot(10:1)
```

You can enable figure captions by `fig_caption: yes` in YAML:

    output:
      rmarkdown::html_vignette:
        fig_caption: yes

Then you can use the chunk option `fig.cap = "Your figure caption."` in **knitr**.

## More Examples

You can write math expressions, e.g. $Y = X\beta + \epsilon$, footnotes^[A footnote here.], and tables, e.g. using `knitr::kable()`.

```{r, echo=FALSE, results='asis'}
knitr::kable(head(mtcars, 10))
```

Also a quote using `>`:

> "He who gives up [code] safety for [code] speed deserves neither."
([via](https://twitter.com/hadleywickham/status/504368538874703872))
